---
title: "Lab 2 Programming in R"
output: pdf_document
---

Part 1:Conditional Statement
Question 1: In a lottery with number from 0 to 100, one will receive an "Apple" with a number less than 30 or greater than 90, or a "Banana" with a number greater than 40, or a "Cat" with the remaining number. Write a lottery_prize function so that with input of the lottery number, an output of prize would be shown. Notice if an out of range number shows, it should provide a warning "Fake Lottery Number." and prize should be 0.
```{r}
lottery_prize <- function(x) {
  if (x < 0 | x > 100) {
      cat("Fake Lottery Number.\n")
      cat(paste("Prize of lottery number", x, "is 0 \n" ))
 } else if (x < 30 | x > 90) {
      cat(paste("Prize of lottery number", x, "is Apple \n"))
 } else if (x > 40) {
      cat(paste("Prize of lottery number", x, "is Banana \n"))
 } else {
        cat(paste("Prize of lottery number", x, "is Cat \n"))
  }
}


```

```{r}
# According to the requirement, if a is sent into lottery_prize(), a>100 or a<0 will result in 0, 0<= a < 30 and 90 <a<=100 result in Apple. 40 < a <=90 results in Banana, and 30 <= a <= 40 results in Cat. Some examples are shown as follows.
lottery_prize(-1)
lottery_prize(45)
lottery_prize(100)
```

Question 2: Load the “hotel_bookings.csv” dataset. Create a binary variable, lead_time01, that contains a 1 if lead_time contains a value above or equal to its median, and a 0 if lead_time contains a value below its median. You need to use the following two ways to create the new column.
(Hint: You can compute the median using the median() function)
```{r}
bookings <- read.csv("hotel_bookings.csv")
median_lead_time <- median(bookings$lead_time)
```

a)	Write a custom function that can be used to check whether the value if above the median first and then use the sapply function to create the new column. 
```{r}
check_function <- function(value, median) {
  if (value >= median) {
    return(1)
  } else {
    return(0)
  }
}

bookings$lead_time01 <- sapply(bookings$lead_time, check_function, median_lead_time)
```

b)	Use the “ifelse” function to create the new column.
```{r}
bookings$lead_time01 <- ifelse(bookings$lead_time >= median_lead_time, 1, 0)
```
Part 2:Loop Expressions
Question 1 Use while loop to calculate value of the sum of even numbers minus the sum of odd numbers from 1 to 100. (Hint: You can calculate the result in this way: -1+2-3+4-5+6+…-97+98-99+100).
```{r}
      x = 1
      sum_P2_q1 <- 0
      indi <- -1 # Here, I use an indicator as sign of the number.
      
      while(x<=100)
        {
        sum_P2_q1 <- sum_P2_q1 + indi * x
        x <- x + 1
        indi <- indi * (-1)
      }
      sum_P2_q1
```

Question 2
a) Write a function factorial() that prints out the x factorial x!=x*(x-1)8(x-2)8...*1(Note that: 0! = 1). And then use the function to calculate 0!, 3! and 7!.
```{r}
      calculate <- function( x ){
        if ( x >= 0){
          i = factorial( x ) 
        }
        else {
          i = print( "error" )
        }
        return( i )
      }
```
check answer
```{r}
      calculate(0)
      calculate(3)
      calculate(7)
```

b) Write a function to calculate the combination of m and n(C(m,n), m>=n), you can use the formula below to get the combination: C(m,n)=m!/(n!*(m-n)!) 
```{r}
      mn <- function( m , n ){ 
        if ( m >= 0){
          if( n >= 0){
            if ( m - n >= 0){
              i = factorial( m ) / ( factorial( n ) * factorial( m - n ) ) 
            }
            else {
              i = print( "error" )
            }
          }
          else i = print( "error" )
        }
        else i = print( "error" )
        return( i )
      }
```

c) Calculate C(5,-1),C(5,1),C(5,2)
```{r}
        mn( 5 , -1 )
        mn( 5 , 1 )
        mn( 5 , 2 )
```

Part 3:Data visualization
In this part, you need to use “ggplot2” package to perform data visualization with the “hotel_bookings.csv” dataset. (Hint: Please make sure you have loaded the “ggplot2” module and the “hotel bookings.csv” dataset)
```{r}
library(ggplot2)
setwd("D:/C_Transfer/Desktop/Zhuoqian_Zhang_WUSTL/COURSES/R/Lab_2")
# If the path has changed, then this path to the folder shall be changed, too.
bookings <- read.csv("hotel_bookings.csv")

```

Question 1 
a) Plot a bar chart of the total booking numbers by hotel type (Variable: "hotel").
b) Plot a bar chart of the total booking numbers by market segment (Variable: "market_segment").
c) Change the two bar charts into pie charts.
```{r}
# a)
library(dplyr)
# Calculate total bookings by hotel type
hotel_type_bookings <- bookings %>%
  group_by(hotel) %>%
  summarize(total_bookings = n())  # Create a data frame containing information about types of hotel and numbers of bookings related.


ggplot(hotel_type_bookings, aes(x = hotel, y = total_bookings)) +
  geom_bar(stat = "identity", fill = "#1F77B4", width = 0.5) +
  labs(x = "Hotel Type", y = "Total Bookings") +
  ggtitle("Total Bookings by Hotel Type") +   # The bar chart's title
  theme(plot.title = element_text(hjust = 0.5)) # Center the title
# b)

# Calculate total bookings by market segment
market_segment_bookings <- bookings %>%
  group_by(market_segment) %>%
  summarize(total_bookings = n()) # Create a data frame containing information about market segment and numbers of bookings related.

# Create bar chart
ggplot(market_segment_bookings, aes(x = market_segment, y = total_bookings)) +
  geom_bar(stat = "identity", fill = "#1F77B4") +
  labs(title = "Total Bookings by Market Segment", x = "Market Segment", y = "Total Bookings") +
  theme(plot.title = element_text(hjust = 0.5))
# c)

# Create pie chart for hotel type
ggplot(hotel_type_bookings, aes(x = "", y = total_bookings, fill = hotel)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start=0) +
  labs(title = "Total Bookings by Hotel Type", fill = "Hotel Type", x = NULL, y = NULL) +
  theme(plot.title = element_text(hjust = 0.5))

# Create pie chart for market segment
ggplot(market_segment_bookings, aes(x = "", y = total_bookings, fill = market_segment)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start=0) +
  labs(title = "Total Bookings by Market Segment", fill = "Market Segment", x = NULL, y = NULL) +
  theme(plot.title = element_text(hjust = 0.5))
```

Question 2
b)	Draw a line chart to show the total booking number of each month by hotel type (Variable: "hotel"). (Hint: Please use the new data "booking_month" and the new y variable "booking_number" to draw the line chart).
```{r}
install.packages("dplyr") # If the package is already installed, this line of code will not be needed.
library(dplyr)
booking <- read.csv("hotel_bookings.csv")
booking_month <- booking %>% 
group_by(arrival_date_month, hotel) %>% 
summarise(booking_number = n(), .groups = "drop") %>% 
arrange(factor(arrival_date_month, levels = month.name))

# Define a named vector that maps month names to integers
month_dict <- c(January = 1, February = 2, March = 3, April = 4, May = 5, June = 6,
                July = 7, August = 8, September = 9, October = 10, November = 11, December = 12)

# Use the dictionary to convert month names to integers
booking_month$arrival_date_month <- month_dict[booking_month$arrival_date_month]
month_names <- month.abb

# As the character expression of months cannot be placed into line chart properly, we have to change the character expression of months into integer numbers first. Here I use 1 to 12 to represent January to December separately.


# Draw the line chart
library(ggplot2)
ggplot(booking_month, aes(x = arrival_date_month, y = booking_number, color = hotel)) +
  geom_line() +
  labs(title = "Total Booking Number by Month and Hotel Type",
       x = "Month",
       y = "Booking Number",
       color = "Hotel Type") +
  scale_x_continuous(breaks = 1:12, labels = month_names) +
  scale_color_manual(values = c("#69b3a2", "#404080")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))



#--------------------------------------------------------------

```

Question 3
a)	Create a box plot to display the relationship between deposit type (Variable: "deposit_type") and booking days in advance (Variable: "lead_time") with the dataset “hotel_bookings.csv”.
   
b)	Create a box plot to display the relationship between cancellation (Variable: "is_canceled") and booking days in advance (Variable: "lead_time") with the dataset “hotel_bookings.csv”.
```{r}
# a)
ggplot(bookings, aes(x = deposit_type, y = lead_time)) +
  geom_boxplot() +
  labs(x = "Deposit Type", y = "Lead Time") +
  ggtitle("Relationship Between Deposit Type and Lead Time") +
  theme(plot.title = element_text(hjust = 0.5))

# b)
# Create a new variable "cancellation" based on "is_canceled", if nothing is done, the ggplot() function cannot recognize two groups out of 01 vector of bookings$is_canceled.
bookings$cancellation <- ifelse(bookings$is_canceled == 0, "Not Canceled", "Canceled") 


ggplot(bookings, aes(x = cancellation, y = lead_time)) +
  geom_boxplot() +
  labs(x = "Cancellation", y = "Lead Time") +
  ggtitle("Relationship Between Cancellation and Lead Time") +
  theme(plot.title = element_text(hjust = 0.5))

```









